REDO Simulating Dwarf Galaxy Evolution with Individual Star Feedback and Chemical Evolution I: Methods and Test Case

### 2017-11-27 main.tex
finished 4.5

Recent cosmological simulations of dwarf galaxies with detailed feedback over a range of dark matter halo masses find that they exhibit large outflows, with mass loading factors on order of 10 - 1000 \citep{Muratov2015}. However, comparable work of idealized dwarf galaxies with detailed feedback and physics models find more modest mass loading factors \citep{Hu2016,Hu2017}. In Figure~\ref{fig:mass_outflow} we present the mass outflow and mass loading rates for our dwarf galaxy as a function of time, computed at five different positions from the galaxy. We follow \citet{Muratov2015} defining the mass outflow rate ($\dot{M}_{\rm outflow}$) at any given radius to be the sum of the outflow rate in all cells in a spherical annulus of width $dL$ centered at that radius. In other words, $\dot{M}_{\rm outflow} = \sum M_{\rm gas} \times v_{\rm r} / dL$. We choose $dL = 0.1~R_{\rm vir}$, or 2.74 kpc.

The total mass outflow rates and mass loading factors at 0.1, 0.25, 0.5, and 1.0 $R_{\rm vir}$ are shown in Figure~\ref{fig:mass_outflow}. We use the 10 Myr averaged SFR to compute the mass loading rates. The outflow rate at 0.1 $R_{\rm vir}$ is high, corresponding to mass loading factors on order of 50 - 100 throughout the simulation time. This quickly declines towards larger radii, however, with effectively zero outflow past the virial radius. \citet{Muratov2015} finds typical mass loading factors at 0.25 R$_{\rm vir}$ on order of 20-40 for galaxies with $v_{c} = 30$ km s$^{-1}$ at low redshift. Our dwarf galaxy exhibits mass loading factors between 4 and 40 at this radius; this is consistent with the results from \citet{Muratov2015}, but perhaps slightly below the typical values.

Interestingly, the v$_{\rm c}$ $\sim$ 30 km s$^{-1}$ halos examined in \citet{Muratov2015} are more massive than the M$_{\rm vir} = 2.5\times 10^9$ M$_{\odot}$ halo examined here by a factor of a few. Using a fit provided in \citet{Muratov2015} to extrapolate and compare at fixed halo mass, one would expect mass loading factors on order of 100 at 0.25 R$_{\rm vir}$ for our dwarf galaxy, a factor of a few higher than what we find. These differences could be attributed to our lack of cosmological evolution in these isolated simulations, but ultimately requires a larger set of dwarf galaxy simulations to make a more robust comparison. We note, however, that our results are much closer to the \citep{Muratov2015} results than those in \citet{Hu2016,Hu2017}, which find mass loading factors at 0.05 $R_{\rm vir}$ between 1 and 10 for a dwarf galaxy with M$_{\rm vir} = 10^{10}$ M$_{\odot}$, and therefore certainly much smaller mass loading factors at 0.25 $R_{\rm vir}$.

% figure here of mass loading and mass outflow rates
\begin{figure*}
\centering
\includegraphics[width=0.45\linewidth]{total_mass_outflow}
\includegraphics[width=0.45\linewidth]{total_mass_loading}
\caption{Spherical mass outflow rates and mass loading rates over time at 4 different radii from the galaxy.}
\label{fig:mass_outflow}
\end{figure*}

Detailed outflow properties, beyond outflow rates and mass loading factors, are important discriminators between the model dependent feedback physics included in galaxy simulations. \textbf{(do I need an example of this in recent work?)}. In Figure~\ref{fig:outflow_velocity} we present radial velocity distributions of all material outside our dwarf galaxy's disk, and within the halo, broken into three gas phases. Half of the outflow, by mass, is at velocities below 40 km s$^{-1}$, with 75\% at velocities of 70 km s$^{1}$ or below. This gas is dominated by the two warm gas phases, with the WNM dominating at lower velocities, and the WIM at higher velocities. The distribution has a very long tail out to and beyond 500 km s$^{-1}$. The dominant launching mechanism in this simulation are supernovae (Emerick et. al. in prep), generating a rapidly moving and volume filling HIM, consistent with the results in \citet{Hu2016,Hu2017}. However, as shown, the HIM comprises very little of the outflow by mass. Most of the outflowing gas comes from the warm phase, pushed out by the high pressure, fast moving HIM. Some of this warm gas certainly originates from adiabatically and radiatively cooled HIM, however. The amount of transfer between phases in the halo of our galaxy will be investigated in future work.

\begin{figure}
\centering
\includegraphics[width=0.95\linewidth]{outflow_velocity_distribution}
\caption{The time averaged radial velocity distribution of outflowing material external to the disk and within the virial radius of our dwarf galaxy. The outflowing material is multiphase, broken into WNM, WIM, and HIM. See XXX for definitions of these regimes. We note that WNM is often labeled simply as ``cold'', or gas below $T = 10^{4}$K. There is little to no outflowing mass in the CNM or molecular phases.}
\label{fig:outflow_velocity}
\end{figure}

Dwarf galaxies efficiently, and preferentially, eject metals released in stellar feedback from their shallow potential wells \citep{MacLowFerrara1999,FerraraTolstoy2000}. This has been better quantified recently both observationally \citep[e.g.][]{Zahid2012,Peeples2014,McQuinn2015} and with more detailed cosmological simulations \citep{Angles-Alcazar2017,Muratov2017}. In Figure~\ref{fig:outflow_metallicity} we show the average ISM metallicity (black), halo metallicity (blue), and outflow metallicities (orange) at 0.1 (solid), 0.25 (dashed), and 1.0 (dotted) $R_{\rm vir}$ over the simulation time. The ISM metallicity evolves minimally from the initial conditions in this time, while the halo and outflow properties change significantly. It is apparent from this figure that the outflows are preferentially metal enriched, as the halo increases in metallicity by a factor of $\sim 2$ in the simulation time. Although most of gas mass in the halo (outside the disk) remains is within 0.2 $R_{\rm vir}$ (see Section~\ref{sec:obs_metals}), the mass that does make it out far and out side the halo is metal enriched by a factor of $\sim$5-6 compared to the ISM. This is in contrast with the results from \citet{Muratov2017}, which generally shows $Z_{\rm outflow} \simeq Z_{\rm ISM}$ at $0.25~R_{\rm vir}$ and $Z_{\rm outflow} < Z_{\rm ISM}$ at $1.0~R_{\rm vir}$, though both with some scatter that reaches above $Z_{\rm outflow} > Z_{\rm ISM}$, especially at high redshift. However, we note that our simulations certainly do not contain a cosmologically realistic CGM and do not include ongoing cosmological accretion of pristine gas; this will certainly lower the typical halo and outflow metallicities, possibly more in agreement with the findings of \citet{Muratov2017}.

>>>>

Recent cosmological simulations of dwarf galaxies with 
%mm detailed 
feedback over a range of dark matter halo masses find that they exhibit large outflows, with mass loading factors on order of 100--1000 \citep{Muratov2015}. However, comparable models of idealized dwarf galaxies with detailed feedback and physics treatments find more modest mass loading factors \citep{Hu2016,Hu2017}. In Figure~\ref{fig:mass_outflow} we present the mass outflow and mass loading rates for our dwarf galaxy as a function of time, computed at five different positions from the galaxy. We follow \citet{Muratov2015} defining the mass outflow rate ($\dot{M}_{\rm outflow}$) at any given radius to be the sum of the outflow rate in all cells in a spherical annulus of width $dL$ centered at that radius,
\begin{equation} \label{eq:dotM}
\dot{M}_{\rm outflow} = \sum M_{\rm gas} \times v_{\rm r} / dL.
\end{equation} 
We choose $dL = 0.1~R_{\rm vir}$, or 2.74 kpc.

The total mass outflow rates and mass loading factors at 0.1, 0.25, 0.5, and 1.0 $R_{\rm vir}$ are shown in Figure~\ref{fig:mass_outflow}. We use the 10 Myr averaged SFR to compute the mass loading rates. The outflow rate at 0.1 $R_{\rm vir}$ is high, corresponding to mass loading factors on order of 50--100 throughout the simulation time. This quickly declines towards larger radii, however, with effectively zero outflow past the virial radius. \citet{Muratov2015} finds typical mass loading factors at 0.25 R$_{\rm vir}$ on order of 20--40 for galaxies with $v_{c} = 30$ km s$^{-1}$ at low redshift. Our dwarf galaxy exhibits mass loading factors between 4 and 40 at this radius; this is consistent with the results from \citet{Muratov2015}, but perhaps slightly below their typical values.

Interestingly, the v$_{\rm c}\sim$ 30 km s$^{-1}$ halos examined in \citet{Muratov2015} are more massive than the M$_{\rm vir} = 2.5\times 10^9$ M$_{\odot}$ halo examined here by a factor of a few. Using a fit provided in \citet{Muratov2015} to extrapolate and compare at fixed halo mass, one would expect mass loading factors on order of 100 at 0.25 R$_{\rm vir}$ for our dwarf galaxy, a factor of a few higher than what we find. These differences could be attributed to our lack of cosmological evolution in these isolated simulations, but ultimately requires a larger set of dwarf galaxy simulations to make a more robust comparison. We note, however, that our results are much closer to the \citep{Muratov2015} results than those in \citet{Hu2016,Hu2017}, which find mass loading factors at 0.05 $R_{\rm vir}$ between 1 and 10 for a dwarf galaxy with M$_{\rm vir} = 10^{10}$ M$_{\odot}$, and therefore certainly much smaller mass loading factors at 0.25 $R_{\rm vir}$.

% figure here of mass loading and mass outflow rates
\begin{figure*}
\centering
\includegraphics[width=0.45\linewidth]{total_mass_outflow}
\includegraphics[width=0.45\linewidth]{total_mass_loading}
\caption{Spherical mass outflow rates (Eq.~[\ref{eq:dotM}]) and mass loading rates over time at 4 different radii from the galaxy.}
\label{fig:mass_outflow}
\end{figure*}

Detailed outflow properties, beyond outflow rates and mass loading factors, are important discriminators between the model dependent feedback physics included in galaxy simulations. 
%mm \textbf{(do I need an example of this in recent work?)}. [maybe soften to "can help discriminate?"]
In Figure~\ref{fig:outflow_velocity} we present radial velocity distributions of all material outside our dwarf galaxy's disk, and within the halo, broken into three gas phases. Half of the outflow, by mass, is at velocities below 40 km s$^{-1}$, with 75\% at velocities of 70 km s$^{-1}$ or below. This gas is dominated by the two warm gas phases, with the WNM dominating at lower velocities, and the WIM at higher velocities. The distribution has a very long tail out to and beyond 500~km~s$^{-1}$. The dominant launching mechanism in this simulation are supernovae (Emerick et. al. in prep), generating a rapidly moving and volume-filling HIM, consistent with the results in \citet{Hu2016,Hu2017}. However, as shown, the HIM comprises very little of the outflow by mass. Most of the outflowing gas comes from the warm phase, pushed out by the high pressure, fast moving HIM. Some of this warm gas certainly originates from adiabatically and radiatively cooled HIM, however. The amount of transfer between phases in the halo of our galaxy will be investigated in future work.

\begin{figure}
\centering
\includegraphics[width=0.95\linewidth]{outflow_velocity_distribution}
\caption{The time averaged radial velocity distribution of outflowing material external to the disk and within the virial radius of our dwarf galaxy. The outflowing material is multiphase, broken into WNM, WIM, and HIM. See Section~\ref{sec:phase} for definitions of these regimes. We note that WNM is often labeled simply as ``cold'', or gas below $T = 10^{4}$K. There is little to no outflowing mass in the CNM or molecular phases.}
\label{fig:outflow_velocity}
\end{figure}

Dwarf galaxies efficiently, and preferentially, eject metals released in stellar feedback from their shallow potential wells \citep{MacLowFerrara1999,FerraraTolstoy2000}. This has been better quantified recently both observationally \citep[e.g.][]{Zahid2012,Peeples2014,McQuinn2015} and with more detailed cosmological simulations \citep{Angles-Alcazar2017,Muratov2017}. In Figure~\ref{fig:outflow_metallicity} we show the average ISM metallicity (black), halo metallicity (blue), and outflow metallicities (orange) at 0.1 (solid), 0.25 (dashed), and 1.0 (dotted) $R_{\rm vir}$ over the simulation time. The ISM metallicity evolves minimally from the initial conditions in this time, while the halo and outflow properties change significantly. It is apparent from this figure that the outflows are preferentially metal enriched, as the halo increases in metallicity by a factor of $\sim 2$ in the simulation time. Although most of the gas mass in the halo (outside the disk) remains within 0.2 $R_{\rm vir}$ (see Section~\ref{sec:obs_metals}), the mass that does make it 
%mm out far and out side the halo
   to the outer reaches of the halo or beyond 
is metal enriched by a factor of $\sim$5--6 compared to the ISM. This is in contrast with the results from \citet{Muratov2017}, which generally show $Z_{\rm outflow} \simeq Z_{\rm ISM}$ at $0.25~R_{\rm vir}$ and $Z_{\rm outflow} < Z_{\rm ISM}$ at $1.0~R_{\rm vir}$, though both with some scatter that reaches 
%mm above 
$Z_{\rm outflow} > Z_{\rm ISM}$, especially at high redshift. However, we note that our simulations certainly do not contain a cosmologically realistic circumgalactic medium, nor do they include ongoing cosmological accretion of pristine gas; this will certainly lower the typical halo and outflow metallicities, possibly 
%mm more in 
    resulting in better
agreement with the findings of \citet{Muratov2017}.
